Resilience and Home-Space for WSTS \thanks{This work was partly done while the authors were supported by the Agence Nationale de la Recherche grant BraVAS (ANR-17-CE40-0028).}

### 2023-10-23 VMCAI-revised-version/final.tex
effectivity

@@

\noindent
Let us now recall the (most general) definition of well-structured transition systems.
\begin{definition}(Finkel \cite{DBLP:journals/iandc/Finkel90})
A {\em Well-Structured Transition System} (WSTS)  $\mathscr{S}=(S, \rightarrow, \leq)$
is a transition system $(S, \rightarrow)$
equipped with a wqo ${\leq} \subseteq S \times S$ such that   
the transition relation $ \rightarrow$ is (upward) compatible with $\leq$, i.e., for all 
$s_1, t_1 , s_2 \in S$
	with $s_1 \leq s_2$  and $s_1 \rightarrow t_1$, there exists 
	$t_2 \in S$ with 
	$t_1 \leq t_2$ and $s_2 \rightarrow^{*} t_2$.
\end{definition}

@@

Another type of WSTS exists that enjoys downward compatibility.

\begin{definition}(Finkel and Schnoebelen \cite{DBLP:journals/tcs/FinkelS01})
A {\em downward (compatible) well-structured transition system} (downward-WSTS)  $\mathscr{S}=(S, \rightarrow, \leq)$
is a transition system $(S, \rightarrow)$
equipped with a wqo ${\leq} \subseteq S \times S$ such that   
the transition relation $ \rightarrow$ is downward compatible with $\leq$, i.e., for all 
$s_1, t_1 , s_2 \in S$
	with $s_2 \leq s_1$  and $s_1 \rightarrow t_1$, there exists 
	$t_2 \in S$ with 
	$t_2 \leq t_1$ and $s_2 \rightarrow^{*} t_2$.
\end{definition}

@@

%An equivalent formulation is : Does there exist $t \in S$ such that $s_0 \longrightarrow^{*} t$ and $s \leq t$ ?
With the $\pred$-basis-effective hypothesis, we obtain:

\begin{theorem}(Finkel and Schnoebelen  \cite{DBLP:journals/tcs/FinkelS01}, Abdulla and al. \cite{DBLP:journals/iandc/AbdullaCJT00})
 A finite basis of $ \pred^*(U)$ is computable for any $\pred$-basis-effective WSTS $\mathscr{S}=(S, \rightarrow, \leq)$ and any upward-closed set $U \subseteq S$ given with its finite basis $B_U$. Hence coverability is decidable.
\end{theorem}

With the $\post$-ideal-effective hypothesis, we obtain the decidability of coverability for WBTS (without the $\pred$-basis-effective hypothesis).

\begin{theorem}(Blondin and al. ~\cite{DBLP:journals/lmcs/BlondinFM17})
Coverability is decidable for any $\post$-ideal-effective WBTS.
\end{theorem}

Downward-WSTS enjoy a powerfull property.

\begin{theorem}(Finkel and Schnoebelen \cite{DBLP:journals/tcs/FinkelS01})
$\post$-effective finitely branching downward-WSTS with reflexive compatibility are $\post^*$-basis-effective.
\end{theorem}
%
%Coverability is  decidable for a class of WSTS without the effective $\pred$-basis hypothesis.
%
%\begin{theorem} \cite{BFM-ic17}
%Coverability is decidable for ideally-effective WSTS.
%\end{theorem}

@@

On the positive side, let us recall a result about {\sc Bounded-state-resilience} (called resilience in \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}).
%
\begin{theorem}\label{ref ozkan}
(\"Ozkan~\cite{DBLP:conf/gg/Ozkan22}, \"Ozkan and W{\"{u}}rdemann \cite{DBLP:journals/corr/abs-2108-00889})
{\sc Bounded-state-resilience} and {\sc $k$-state-resilience} are decidable for $\pred$-basis-effective, $\post^*$-basis-effective WSTS with strong compatibility
when
$\Safe=\mathop{\uparrow} \Safe$.
\end{theorem}

@@

\begin{proposition}[Proposition 1 in \cite{DBLP:conf/gg/Ozkan22}]\label{post*}
For $\post$-effective, $\pred$-basis-effective finite-branching WSTS, a finite basis of $\mathop{\uparrow} \post^*(s)$ is computable for every state $s$ iff the downward-reachability problem is decidable.
\end{proposition}

>>>>

\alain{j'ai enlevé les noms lorsqu'on cite un resultat}

@@

\noindent
Let us now recall the (most general) definition of well-structured transition systems.
\begin{definition}[Definition 3.10, \cite{DBLP:journals/iandc/Finkel90}]
A {\em Well-Structured Transition System} (WSTS)  $\mathscr{S}=(S, \rightarrow, \leq)$
is a transition system $(S, \rightarrow)$
equipped with a wqo ${\leq} \subseteq S \times S$ such that   
the transition relation $ \rightarrow$ is (upward) compatible with $\leq$, i.e., for all 
$s_1, t_1 , s_2 \in S$
	with $s_1 \leq s_2$  and $s_1 \rightarrow t_1$, there exists 
	$t_2 \in S$ with 
	$t_1 \leq t_2$ and $s_2 \rightarrow^{*} t_2$.
\end{definition}

@@

Another type of WSTS exists that enjoys downward compatibility.
%
%		Finkel and Schnoebelen, 
%
\begin{definition}[Definition 5.1, \cite{DBLP:journals/tcs/FinkelS01}]
A {\em downward (compatible) well-structured transition system} (downward-WSTS)  $\mathscr{S}=(S, \rightarrow, \leq)$
is a transition system $(S, \rightarrow)$
equipped with a wqo ${\leq} \subseteq S \times S$ such that   
the transition relation $ \rightarrow$ is downward compatible with $\leq$, i.e., for all 
$s_1, t_1 , s_2 \in S$
	with $s_2 \leq s_1$  and $s_1 \rightarrow t_1$, there exists 
	$t_2 \in S$ with 
	$t_2 \leq t_1$ and $s_2 \rightarrow^{*} t_2$.
\end{definition}

@@

%An equivalent formulation is : Does there exist $t \in S$ such that $s_0 \longrightarrow^{*} t$ and $s \leq t$ ?
With the $\pred$-basis-effective hypothesis, we obtain:
%
%			(Finkel and Schnoebelen, Abdulla and al., 
%
\begin{theorem}[Theorem 3.6, \cite{DBLP:journals/tcs/FinkelS01}, Theorem 4.1, \cite{DBLP:journals/iandc/AbdullaCJT00}]
 A finite basis of $ \pred^*(U)$ is computable for any $\pred$-basis-effective WSTS $\mathscr{S}=(S, \rightarrow, \leq)$ and any upward-closed set $U \subseteq S$ given with its finite basis $B_U$. Hence coverability is decidable.
\end{theorem}

With the $\post$-ideal-effective hypothesis, we obtain the decidability of coverability for WBTS (without the $\pred$-basis-effective hypothesis).
%
%		(Blondin and al., 
%
\begin{theorem}[Corollaire 4.4, ~\cite{DBLP:journals/lmcs/BlondinFM17}]
Coverability is decidable for any $\post$-ideal-effective WBTS.
\end{theorem}

Downward-WSTS enjoy a powerfull property.
%
%		Finkel and Schnoebelen, 
%
\begin{theorem}[Proposition 5.4, \cite{DBLP:journals/tcs/FinkelS01}]
Finitely branching downward-WSTS with reflexive compatibility and $\post$-effective are $\post^*$-basis-effective.
\end{theorem}
%
%Coverability is  decidable for a class of WSTS without the effective $\pred$-basis hypothesis.
%
%\begin{theorem} \cite{BFM-ic17}
%Coverability is decidable for ideally-effective WSTS.
%\end{theorem}

@@

On the positive side, let us recall a result about {\sc Bounded-state-resilience} (called resilience in \cite{DBLP:conf/gg/Ozkan22,DBLP:journals/corr/abs-2108-00889}).
%
%		(\"Ozkan,   \"Ozkan and W{\"{u}}rdemann,  
%
\alain{je ne comprends pas pourquoi les citations ne sortent pas comme les autres}
\begin{theorem}\label{ref ozkan}[Theorem 1,~\cite{DBLP:conf/gg/Ozkan22}, Theorem 1, ~\cite{DBLP:journals/corr/abs-2108-00889}]
{\sc Bounded-state-resilience} and {\sc $k$-state-resilience} are decidable for $\pred$-basis-effective, $\post^*$-basis-effective WSTS with strong compatibility
when
$\Safe=\mathop{\uparrow} \Safe$.
\end{theorem}

@@

\begin{proposition}[Proposition 1, \cite{DBLP:conf/gg/Ozkan22}]\label{post*}
For $\post$-effective, $\pred$-basis-effective finite-branching WSTS, a finite basis of $\mathop{\uparrow} \post^*(s)$ is computable for every state $s$ iff the downward-reachability problem is decidable.
\end{proposition}
